Inefficient Concessions and Mediation

### 2016-09-27 draft.tex
remove references to 'formal'

\title{Inefficient Concessions and Mediation} 
\author{Kristy Buzard\\
Syracuse University \\
kbuzard@syr.edu\\
Ben Horne\\ 
University of California, San Diego}
%To Do : define what is ineffieicncy

@@

This paper has three parts. First, it explores the role of costly concessions in a game theoretic framework, evoking a two-sided version of Spence signaling (1973). Second, after considering existing explanations for inefficient concessions, it proposes a new theoretical explanation for why inefficiency might occur in a conflict setting: inefficient concessions may be preferred if the most efficient concessions are seen as having additional, future costs.\footnote{The type of inefficient concessions addressed here are those whose value to the recipient are lower than the cost to the giver.} The Golan Heights, for example, will not be conceded if Israel foresees future security risks that outweigh the immediate benefits of peace. Finally, this paper uses a mechanism design framework to model a third party's role in bringing about more efficient concessions. Many scholars contend that mediators with enforcement capabilities can help to resolve conflict; this formal analysis provides an explanation for why enforcement is effective.

@@

\section{Formal model description} %A DIFFERENT MODEL BUT SAME RESULT AS FR 10; THIS SHOULD BUILD OUR CONFIDENCE IN MODELLING. EMPHASIZE SCENARION (EGYPT USRAEL ETC) THAT ARE BETTER TO MODEL WITH Our MODEL. IT CAN ALSO BE THOUGHT OF AS REDUCED FORM OF A BARGAINING MOEL, BUT IT DOES NOT NEED TO BE! NEED TO GIVE EXAMPLES WHERE Our MODEL IS BETTER THAN BARGAINING- EXAMPLES. WHEN TRUST IS AN ISSUE (PUT THIS IN THE TRUST SECTION)
%Need to add risk aversion in MAIN TEXT

@@

%\captionof{table}{Stage Game Payoffs}
\begin{center}
\begin{tabular}{|l|c|c|}
  \hline      & Trust & Fight \\ \hline
	 Trust& T, T& -D, T+W \\ \hline
	Fight & T+W,-D& W-D, W-D \\ \hline
\end{tabular}
\end{center}

@@

\section{Analysis of benchmark cases }
%examples!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!

When countries are faced with doubts about each other's motives, they have several options of how to proceed. There are three basic types of equilibria that this analysis will examine.\footnote{If the discount factor for the high type ($\delta_h$) exceeds the cutoff found in expression (1), there are many types of equilibria; consider for instance oscillating every other period between Peace and War with a grim trigger War threat. In this paper, we are not interested in these equilibria because they are not as realistic for applications to conflict scenarios. These are also always payoff dominated by at least one of the other equilibria discussed, so are not attractive.} A thorough analysis of the different types of equilibria, as well as all proofs for this section, can be found in the appendix. 
\\
\\
{\bf \emph{Pooling equilibrium}\\}
\vskip-.2in
In the \emph{pooling equilibrium}, both countries do not give concessions in period 0, then always play Fight in every period beginning with period 1. 
Countries continue to doubt, doing nothing to bridge the credibility gap, and thus not cooperating. This is a \emph{pooling equilibrium}, and is likely to happen when countries are in long-set patterns of distrust. Active fighting is not necessary; remaining outside of a state of peace is all that is required.  See Lemma 1 and Lemma 2 in the appendix for a formal characterization of this type of equilibrium.

@@

In period 0, after receiving his concession, a country makes the decision to allocate it between military and civil society. $\alpha_i$ is the portion of his received concession that country i chooses to dedicate to civil society; $(1- \alpha_i$) is the portion country i dedicates to military buildup. This decision is a simple optimization problem for each country. Parameters are common knowledge so the result of this decision is well known should the country's type be known. Concessions are immediately converted into civil society or military gifts and incorporated into payoff functions. 
 $C_h(\cdot)=C_l(\cdot)=1$ as earlier in the model description. Periods 1 to $\infty$ are as before but with the following stage game payoffs:
\newpage
%\captionof{table}{Modified Stage Game Payoffs}
\begin{center}

\begin{tabular}{|l|c|c|}
  \hline      & Trust & Fight \\ \hline
	 Trust& \small$T(s_2+\alpha_2g_1)$, & \small$-D(m_2+(1-\alpha_2)g_1),$ \\
  & \small$T(s_1+\alpha_1g_2)$ & \small$T(s_1+\alpha_1g_2)  +W(m_2+(1-\alpha_2)g_1) $\\ \hline
	Fight & \small$T(s_2+\alpha_2g_1)+W(m_1+(1-\alpha_1)g_2),$& \small$W(m_1+(1-\alpha_1)g_2)-D(m_2+(1-\alpha_2)g_1),$ \\
	 & \small$-D(m_1+(1-\alpha_1)g_2) $ & \small$W(m_2+(1-\alpha_2)g_1)-D(m_1+(1-\alpha_1)g_2)$ \\ \hline
\end{tabular}
\end{center}
\vskip.1in
{\bf Theorem 4:} Under some parameters, the optimal equilibrium uses inefficient concessions. \\
 \emph{Proof}:   Because it is relatively complicated to calculate equilibrium concessions in the general case, a slight simplification of the general model will be made for proof of inefficient concessions. Restrict functional forms of $T(\cdot),W(\cdot)$ and $D(\cdot)$ so that each is a scalar multiplied by its arguments. Also assume that $s_1=s_2=m_1=m_2=1.$ Because of linearity, equilibrium optimization has high types choosing $\alpha=1$ (because a concession would only come from another high type in equilibrium) and low types choosing $\alpha=0$. Note that none of these simplifications is essential for the results to hold but we make them for ease of calculation and demonstration of the results. 
The low type's IC constraint binds so with $e=1$, $g^*$ can be calculated to be $g^*= \frac{p(T+D)}{1-pT}$.
Under an equilibrium with purely inefficient concessions $(e=0)$, the minimum separating concession would be $g'=p(T+D)$.

@@

\section{Mediation}
All results in this paper thus far have relied on bilateral engagement. We will now consider the involvement of a mediator which will be modeled as a mechanism.  Mediation has been studied to some degree in the formal literature and has most recently been modeled using mechanism design. Applying mechanism design to mediation has been used, to date, by Fey and Ramsay (2008, 2010 and 2011), Bester and Warneryd (2006) and Horner et. al. (2010).

@@

$M$ inputs the reports of types $t_1$ and $t_2$ and outputs the required concessions $g_1$ and $g_2$. Formally, $ M:(t_1,t_2)\rightarrow( g_1(\cdot), g_2(\cdot))$.  The countries are bound to send the mandated concessions. The benchmark case requires the countries to send the same concession to both types, since types are private information.  $M$ can differentiate based on recipient's type. Otherwise, the setup is identical to the benchmark case.

@@

The effectiveness of mediation is debated in the literature; this paper shows that a third party mediator can remedy this inefficiency if she is trusted. One result of this formal exposition is to show the mechanisms through which a mediator can help. Since the mediator eliminates inefficient concessions, the observation that inefficient concession may not do the trick only further indicates the value of a mediator even more.
%our model can also be used to achieve a similar matching result as Prendergast and Stole (2001) who show how if player values his own type, inefficiency can be preferred.

@@

 Let us examine the conditions for a separating equilibrium held in place by a grim trigger.\\
 \\
{\bf Theorem 1:} The strategy of not giving concessions in period 0, then playing Trust unless the other country defects to Fight is an equilibrium strategy for high types if $\delta_h \geq \frac{W}{(1-p)W + p(T+D)}$.\\
 \emph{Proof}:
Countries with the equilibrium strategy to play Trust in the first round reveal themselves as a high type as per Lemma 2. This will be incentive compatible for high types if the expected payoffs from this strategy are greater than the payoffs from deviating against someone playing the pooling (always Fight) strategy. If both countries play Trust in a period, a Peace equilibrium takes place and both countries continue to play peace in every period with the grim trigger threat of the War equilibrium. Low types play Fight in the first period, so a high type that observed Fight being played would respond by playing Fight in period 2 and in all future periods.  If the potential advantages of Peace are large enough to outweigh the risks of encountering a low type in expected utility terms, a high type will be given incentives to play Trust in the first round rather than Fight. Formally, to play Trust, the following must be satisfied:%Make clearer language!!!!!!!
%FUTURE WORK ON DELY OF PEACE- IF PEACE WILL E THERE, YOU MIGHT BE VERY RISK AVERSE TO GETTING SCREWED BY THE LOW TYPE SO KEEP PUTTING OFF THE RISKY FIRST TRUST ROUND
\be
p(\frac{T}{1- \delta_h}) +(1-p)({-D}+\frac{\delta_h(W-D)}{1- \delta_h} )\geq p (T + W + \frac{\de_h}{1-\de_h}(W-D)) +(1-p)\frac{W-D}{1- \delta_h}
\ee

>>>>

\title{Inefficient Concessions and Mediation} 
\author{Kristy Buzard\\
Syracuse University \\
kbuzard@syr.edu\\
\\
Ben Horne\\ 
University of California, San Diego\\
\\}
%To Do : define what is ineffieicncy

@@

This paper has three parts. First, it explores the role of costly concessions in a game theoretic framework, evoking a two-sided version of Spence signaling (1973). Second, after considering existing explanations for inefficient concessions, it proposes a new theoretical explanation for why inefficiency might occur in a conflict setting: inefficient concessions may be preferred if the most efficient concessions are seen as having additional, future costs.\footnote{The type of inefficient concessions addressed here are those whose value to the recipient are lower than the cost to the giver.} The Golan Heights, for example, will not be conceded if Israel foresees future security risks that outweigh the immediate benefits of peace. Finally, this paper uses a mechanism design framework to model a third party's role in bringing about more efficient concessions. Many scholars contend that mediators with enforcement capabilities can help to resolve conflict; this analysis provides an explanation for why enforcement is effective.

@@

\section{Model} %A DIFFERENT MODEL BUT SAME RESULT AS FR 10; THIS SHOULD BUILD OUR CONFIDENCE IN MODELLING. EMPHASIZE SCENARION (EGYPT USRAEL ETC) THAT ARE BETTER TO MODEL WITH Our MODEL. IT CAN ALSO BE THOUGHT OF AS REDUCED FORM OF A BARGAINING MOEL, BUT IT DOES NOT NEED TO BE! NEED TO GIVE EXAMPLES WHERE Our MODEL IS BETTER THAN BARGAINING- EXAMPLES. WHEN TRUST IS AN ISSUE (PUT THIS IN THE TRUST SECTION)
%Need to add risk aversion in MAIN TEXT

@@

\begin{center}
\captionof{table}{Stage Game Payoffs}
\begin{tabular}{|l|c|c|}
  \hline      & Trust & Fight \\ \hline
	 Trust& T, T& -D, T+W \\ \hline
	Fight & T+W,-D& W-D, W-D \\ \hline
\end{tabular}
\end{center}

@@

\section{Analysis of benchmark cases }
When countries are faced with doubts about each other's motives, they have several options of how to proceed. There are three basic types of equilibria that this analysis will examine.\footnote{If the discount factor for the high type ($\delta_h$) exceeds the cutoff found in expression (1), there are many types of equilibria; consider for instance oscillating every other period between Peace and War with a grim trigger War threat. In this paper, we are not interested in these equilibria because they are not as realistic for applications to conflict scenarios. These are also always payoff dominated by at least one of the other equilibria discussed, so are not attractive.} A thorough analysis of the different types of equilibria, as well as all proofs for this section, can be found in the appendix. 
\\
\\
{\bf \emph{Pooling equilibrium}\\}
\vskip-.2in
In the \emph{pooling equilibrium}, both countries do not give concessions in period 0, then always play Fight in every period beginning with period 1. 
Countries continue to doubt, doing nothing to bridge the credibility gap, and thus not cooperating. This is a \emph{pooling equilibrium}, and is likely to happen when countries are in long-set patterns of distrust. Active fighting is not necessary; remaining outside of a state of peace is all that is required.  See Lemma 1 and Lemma 2 in the appendix for a formal characterization of this type of equilibrium.

@@

In period 0, after receiving his concession, a country makes the decision to allocate it between military and civil society. $\alpha_i$ is the portion of his received concession that country i chooses to dedicate to civil society; $(1- \alpha_i$) is the portion country i dedicates to military buildup. This decision is a simple optimization problem for each country. Parameters are common knowledge so the result of this decision is well known should the country's type be known. Concessions are immediately converted into civil society or military gifts and incorporated into payoff functions. 
 $C_h(\cdot)=C_l(\cdot)=1$ as earlier in the model description. Periods 1 to $\infty$ are as before but with the following stage game payoffs:

\begin{center}
\captionof{table}{Modified Stage Game Payoffs}
\begin{tabular}{|l|c|c|}
  \hline      & Trust & Fight \\ \hline
	 Trust& \small$T(s_2+\alpha_2g_1)$, & \small$-D(m_2+(1-\alpha_2)g_1),$ \\
  & \small$T(s_1+\alpha_1g_2)$ & \small$T(s_1+\alpha_1g_2)  +W(m_2+(1-\alpha_2)g_1) $\\ \hline
	Fight & \small$T(s_2+\alpha_2g_1)+W(m_1+(1-\alpha_1)g_2),$& \small$W(m_1+(1-\alpha_1)g_2)-D(m_2+(1-\alpha_2)g_1),$ \\
	 & \small$-D(m_1+(1-\alpha_1)g_2) $ & \small$W(m_2+(1-\alpha_2)g_1)-D(m_1+(1-\alpha_1)g_2)$ \\ \hline
\end{tabular}
\end{center}
\vskip.1in
{\bf Theorem 4:} Under some parameters, the optimal equilibrium uses inefficient concessions. \\
 \emph{Proof}:   Because it is relatively complicated to calculate equilibrium concessions in the general case, a slight simplification of the general model will be made for proof of inefficient concessions. Restrict functional forms of $T(\cdot),W(\cdot)$ and $D(\cdot)$ so that each is a scalar multiplied by its arguments. Also assume that $s_1=s_2=m_1=m_2=1.$ Because of linearity, equilibrium optimization has high types choosing $\alpha=1$ (because a concession would only come from another high type in equilibrium) and low types choosing $\alpha=0$. Note that none of these simplifications is essential for the results to hold but we make them for ease of calculation and demonstration of the results. 
The low type's IC constraint binds so with $e=1$, $g^*$ can be calculated to be $g^*= \frac{p(T+D)}{1-pT}$.
Under an equilibrium with purely inefficient concessions $(e=0)$, the minimum separating concession would be $g'=p(T+D)$.

@@

\section{Mediation}
All results in this paper thus far have relied on bilateral engagement. We will now consider the involvement of a mediator which will be modeled as a mechanism. In the theoretical literature, mediation has most recently been modeled using mechanism design (Fey and Ramsay 2008, 2010 and 2011; Bester and Warneryd 2006; and Horner et. al. 2010).

@@

$M$ inputs the reports of types $t_1$ and $t_2$ and outputs the required concessions $g_1$ and $g_2$ so that $ M:(t_1,t_2)\rightarrow( g_1(\cdot), g_2(\cdot))$.  The countries are bound to send the mandated concessions. The benchmark case requires the countries to send the same concession to both types, since types are private information.  $M$ can differentiate based on recipient's type. Otherwise, the setup is identical to the benchmark case.

@@

The effectiveness of mediation is debated in the literature; this paper shows that a third party mediator can remedy this inefficiency if she is trusted. One contribution of this paper is to show the mechanisms through which a mediator can help. Since the mediator eliminates inefficient concessions, the observation that inefficient concession may not do the trick only further indicates the value of a mediator even more.
%our model can also be used to achieve a similar matching result as Prendergast and Stole (2001) who show how if player values his own type, inefficiency can be preferred.

@@

 Let us examine the conditions for a separating equilibrium held in place by a grim trigger.\\
 \\
{\bf Theorem 1:} The strategy of not giving concessions in period 0, then playing Trust unless the other country defects to Fight is an equilibrium strategy for high types if $\delta_h \geq \frac{W}{(1-p)W + p(T+D)}$.\\
 \emph{Proof}:
Countries with the equilibrium strategy to play Trust in the first round reveal themselves as a high type as per Lemma 2. This will be incentive compatible for high types if the expected payoffs from this strategy are greater than the payoffs from deviating against someone playing the pooling (always Fight) strategy. If both countries play Trust in a period, a Peace equilibrium takes place and both countries continue to play peace in every period with the grim trigger threat of the War equilibrium. Low types play Fight in the first period, so a high type that observed Fight being played would respond by playing Fight in period 2 and in all future periods.  If the potential advantages of Peace are large enough to outweigh the risks of encountering a low type in expected utility terms, a high type will be given incentives to play Trust in the first round rather than Fight. To play Trust, the following must be satisfied:%Make clearer language!!!!!!!
%FUTURE WORK ON DELY OF PEACE- IF PEACE WILL E THERE, YOU MIGHT BE VERY RISK AVERSE TO GETTING SCREWED BY THE LOW TYPE SO KEEP PUTTING OFF THE RISKY FIRST TRUST ROUND
\be
p(\frac{T}{1- \delta_h}) +(1-p)({-D}+\frac{\delta_h(W-D)}{1- \delta_h} )\geq p (T + W + \frac{\de_h}{1-\de_h}(W-D)) +(1-p)\frac{W-D}{1- \delta_h}
\ee
